Dependent Object Types

### 2012-08-05 article.tex
functions as sugar
basic ex

%% Syntax
% Header for typing rules
\newcommand{\judgement}[2]{{\bf #1} \hfill #2}
% Refinement
\newcommand{\refine}[2]{\left\{#1 \Rightarrow #2 \right\}}
\newcommand{\mlrefine}[2]{\{#1 \Rightarrow #2 \}}
% Field definitions
\newcommand{\ldefs}[1]{\left\{#1\right\}}
\newcommand{\mlldefs}[1]{\{#1\}}
% Member sequences
\newcommand{\seq}[1]{\overline{#1}}
% Lambda
\newcommand{\dabs}[3]{(#1\!:\!#2)\Rightarrow #3}
\newcommand{\abs}[3]{\lambda #1\!:\!#2.#3}
% Application
\newcommand{\app}[2]{#1\;#2}
% Method Application
\newcommand{\mapp}[3]{#1.#2(#3)}
% Substitution
\newcommand{\subst}[3]{[#1/#2]#3}
% Object creation
\newcommand{\new}[3]{\textbf{val }#1 = \textbf{new }#2 ;\; #3}
\newcommand{\mlnew}[3]{\textbf{val }#1 = \textbf{new }#2 ;\;\\&#3}
%\renewcommand{\new}[3]{#1 \leftarrow #2 \,\textbf{in}\, #3}
% Field declaration
\newcommand{\Ldecl}[3]{#1 : #2..#3}%{#1 \operatorname{>:} #2 \operatorname{<:} #3}
\newcommand{\ldecl}[2]{#1 : #2}
\newcommand{\mdecl}[3]{#1 : #2 \tfun #3}
% Top and Bottom
\newcommand{\Top}{\top}%{\textbf{Top}}
\newcommand{\Bot}{\bot}%\textbf{Bot}}
% Environment extension
%\newcommand{\envplus}[1]{\uplus \{ #1 \}}
\newcommand{\envplus}[1]{, #1}
% Reduction
\newcommand{\reduction}[4]{#1 \operatorname{|} #2 \reduces #3 \operatorname{|} #4}

@@

\subsection{Sugar: Functions}

\subsection{Basics: Booleans, Error, \ldots}
This program defines a $\mi{root}$ object with basic types
($\mi{Unit}$, $\mi{Boolean}$, $\mi{Nat}$) and values ($\mi{true}$,
$\mi{false}$, $\mi{error}$, $\mi{zero}$). For brevity, I've omitted
the code related to natural numbers. During the object creation, the
value labels such as $\mi{false}$ are all initialized.
\begin{align*}
&\mlnew {\mi{root}} {\Top \mlrefine r {\\\displaybreak[3]
&\ \Ldecl {\mi{Unit}} \Bot \Top\\
&\ \ldecl {\mi{unit}} {\Top \tfun {r.\mi{Unit}}}\\
&\ \Ldecl {\mi{Boolean}} \Bot {\Top \mlrefine z {\\
&\ \gap \ldecl {\mi{ifNat}} {(r.\mi{Unit} \tfun r.\mi{Nat}) \tfun (r.\mi{Unit} \tfun r.\mi{Nat}) \tfun r.\mi{Nat}}\\
&\gap}}\\
&\ \ldecl {\mi{false}} {{r.\mi{Unit}} \tfun {r.\mi{Boolean}}}\\
&\ \ldecl {\mi{true}} {{r.\mi{Unit}} \tfun {r.\mi{Boolean}}}\\
&\ \ldots\\\displaybreak[3]
&\ \ldecl {\mi{error}} {{r.\mi{Unit}} \tfun \Bot}\\
&}\mlldefs{\\\displaybreak[3]
&\ \mi{unit}  = \abs x \Top {\new u {\mi{root}.\mi{Unit}} u}\\
%
&\ \mi{false} = \abs u {\mi{root}.\mi{Unit}} {\\
&\ \gap \mlnew {\mi{ff}} {\mi{root}.\mi{Boolean} \mlldefs{\\\displaybreak[3]
&\ \gap\gap {\mi{ifNat}} = \abs t {\mi{root}.\mi{Unit} \tfun \mi{root}.\mi{Nat}} {\\&\gap\gap\gap\abs e {\mi{root}.\mi{Unit} \tfun \mi{root}.\mi{Nat}}\\&\gap\gap\gap\app e {\mi{root}.\mi{unit}}}\\
&\ \gap}}{
\ \gap \mi{ff}}
&}\\
%
&\ \mi{true} = \ldots\\
%
&\ \mi{error} = \abs u {\mi{root}.\mi{Unit}} {\app {\mi{root}.\mi{error}} u}\\
&\ \ldots\\
&}}{
\ldots
}
\end{align*}

>>>>

%% Syntax
% Header for typing rules
\newcommand{\judgement}[2]{{\bf #1} \hfill #2}
% Refinement
\newcommand{\refine}[2]{\left\{#1 \Rightarrow #2 \right\}}
\newcommand{\mlrefine}[2]{\{#1 \Rightarrow #2 \}}
% Field definitions
\newcommand{\ldefs}[1]{\left\{#1\right\}}
\newcommand{\mlldefs}[1]{\{#1\}}
% Member sequences
\newcommand{\seq}[1]{\overline{#1}}
% Lambda
\newcommand{\dabs}[3]{(#1\!:\!#2)\Rightarrow #3}
\newcommand{\abs}[3]{\lambda #1\!:\!#2.#3}
% Method Application
\newcommand{\mapp}[3]{#1.#2(#3)}
% Substitution
\newcommand{\subst}[3]{[#1/#2]#3}
% Object creation
\newcommand{\new}[3]{\textbf{val }#1 = \textbf{new }#2 ;\; #3}
\newcommand{\mlnew}[3]{\textbf{val }#1 = \textbf{new }#2 ;\;\\&#3}
%\renewcommand{\new}[3]{#1 \leftarrow #2 \,\textbf{in}\, #3}
% Field declaration
\newcommand{\Ldecl}[3]{#1 : #2..#3}%{#1 \operatorname{>:} #2 \operatorname{<:} #3}
\newcommand{\ldecl}[2]{#1 : #2}
\newcommand{\mdecl}[3]{#1 : #2 \tfun #3}
% Top and Bottom
\newcommand{\Top}{\top}%{\textbf{Top}}
\newcommand{\Bot}{\bot}%\textbf{Bot}}
% Environment extension
%\newcommand{\envplus}[1]{\uplus \{ #1 \}}
\newcommand{\envplus}[1]{, #1}
% Reduction
\newcommand{\reduction}[4]{#1 \operatorname{|} #2 \reduces #3 \operatorname{|} #4}

% Sugar
\newcommand{\arrow}[2]{#1\rightarrow_s#2}
\newcommand{\fun}[4]{\textbf{fun } (#1:#2)\;#3\;#4}
\newcommand{\app}[2]{(\textbf{app }#1\;#2)}
\newcommand{\cast}[2]{(\textbf{cast }#1\;#2)}

@@

\subsection{Sugar: Functions}
Like in Scala, we can encode functions as objects with a special
method. We will freely use the following sugar in the remaining of
this paper. Note that the variable $z$ must be fresh.
\begin{align*}
\arrow S T &\iff {\Top \refine z {\mdecl {\mi{apply}} S T}}\\
\fun x S T t &\iff \new z {\arrow S T} \ldefs{\mi{apply}(x) = t}\\
\app f x &\iff \mapp f {\mi{apply}} x\\
\cast T t&\iff \app {\fun x T T x} t
\end{align*}

In the remaining examples, for brevity, we will also use $\abs x S t$
for $\fun x S \_ t$ where the return type can be somewhat inferred
from the context.

\subsection{Basics: Booleans, Error, \ldots}
This program defines a $\mi{root}$ object with basic types
($\mi{Unit}$, $\mi{Boolean}$, $\mi{Nat}$) and methods ($\mi{true}$,
$\mi{false}$, $\mi{error}$, $\mi{zero}$). For brevity, we've omitted
the code related to natural numbers. During the object creation, the
method labels such as $\mi{false}$ are all initialized.
\begin{align*}
&\mlnew {\mi{root}} {\Top \mlrefine r {\\\displaybreak[3]
&\ \Ldecl {\mi{Unit}} \Bot \Top\\\displaybreak[3]
&\ \mdecl {\mi{unit}} \Top {r.\mi{Unit}}\\\displaybreak[3]
&\ \Ldecl {\mi{Boolean}} \Bot {\Top \mlrefine z {\\\displaybreak[3]
&\ \gap \ldecl {\mi{ifNat}} \arrow {(\arrow {r.\mi{Unit}} {r.\mi{Nat}})} {\arrow {(\arrow {r.\mi{Unit}} {r.\mi{Nat}})} {r.\mi{Nat}}}\\\displaybreak[3]
&\gap}}\\\displaybreak[3]
&\ \mdecl {\mi{false}} {r.\mi{Unit}} {r.\mi{Boolean}}\\\displaybreak[3]
&\ \mdecl {\mi{true}} {r.\mi{Unit}} {r.\mi{Boolean}}\\\displaybreak[3]
&\ \mdecl {\mi{error}} {r.\mi{Unit}} \Bot\\\displaybreak[3]
& \ \ldots\\\displaybreak[3]
&}\mlldefs{\\\displaybreak[3]
&\ \mi{unit}(x) = {\new u {\mi{root}.\mi{Unit}} u}\\\displaybreak[3]
%
&\ \mi{false}(u) =\\\displaybreak[3]
&\ \gap \mlnew {\mi{ff}} {\mi{root}.\mi{Boolean} \mlldefs{\\\displaybreak[3]
&\ \gap\gap {\mi{ifNat}} = \abs t {\arrow {\mi{root}.\mi{Unit}} {\mi{root}.\mi{Nat}}} {\\&\gap\gap\gap\abs e {\arrow {\mi{root}.\mi{Unit}} {\mi{root}.\mi{Nat}}}\\&\gap\gap\gap\app e {\mi{root}.\mi{unit}}}\\\displaybreak[3]
&\ \gap}}{
\ \gap \mi{ff}}\\\displaybreak[3]
%
&\ \mi{true}(u) = \ldots\\\displaybreak[3]
%
&\ \mi{error}(u) = {\app {\mi{root}.\mi{error}} u}\\\displaybreak[3]
&\ \ldots\\\displaybreak[3]
&}}{
\ldots
}
\end{align*}
